 Walkthrough

### 2023-05-09 breakingnewsstory.qmd
final commit

---
title: "Breaking News Story Sample"
author: "Giovanna Lino"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

@@

```{r}

@@

```{r}

@@

nonwhite_voters_average <- inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_race_nonwhite_abovebelow_natl) %>% 
  summarise(average = mean(pct_r_2020))

nonwhite_voters_average

@@

inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise(average = mean(pct_r_2020))
```

#  \# `r nonwhite_voters_average` of Trump's voters are part of the counties above national average for people who identify as nonwhite.

In the aftermath of the 2020 Presidential campaign, both the results of the election and the demographics of the voter pool of each county were analyzed for both the Republican candidate Donald Trump and the Democratic candidate Joe Biden. Particularly, Trump's mean percentage of votes by U.S. counties was analyzed for both the number of nonwhite voters and the number of college graduates and if they were above or below the national U.S. average.

It comes as a surprise that more than `r nonwhite_voters_average` of Trump's voters are part of the counties above the national average for people who identify as nonwhite, while approximately 68% of votes are part of counties below the national average for non white-identified people. Furthermore, on average, `pct_ed_college_all_abovebelow_natl` of Trump's votes come from counties with people above the national average for college graduates, versus 68% below the national average for college graduates.

These results come as a surprise since it is known that the Democratic party has the majority of college graduate and nonwhite votes. Although this is still the case, the Democratic party cannot solely rely on these metrics to ensure victory since more college graduates and nonwhite voters are increasing their support for the Republican party.

This trend can be seen in states such as Texas. Some working-class Hispanics do not feel completely represented by the Democratic party in the last gubernatorial race in November, 2022. The Democratic candidate, Beto O'Rourke is a citizen of El Paso and has been called out by activists from the organization Sunrise El Paso for not supporting them in their fight against the privatization of their electric companies. The city of El Paso is not on the Texas electrical grid, allowing for the company J.P Morgan to privatize it and charge expensive bills. Representatives of Sunrise El Paso claim that O'Rourke is too much of a Central Democrat and does not want to compromise the funding of his campaign by selecting fights with big corporations.

>>>>

---
title: "Breaking News Story"
author: "Giovanna Lino"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

@@

```{r, echo=FALSE}

@@

```{r, echo=FALSE}
# By joining both datasets, I am grouping/summarizing to calculate the mean pct of Trump 2020 vote by counties above or below the national average for people who identify as nonwhite.

#Here is the ABOVE value

suppressMessages({nonwhite_voters_average_above <- inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_race_nonwhite_abovebelow_natl) %>% 
  summarise(average_above_nonwhite = mean(pct_r_2020)) %>% 
  filter(pct_race_nonwhite_abovebelow_natl=="ABOVE") %>% 
  pull(average_above_nonwhite)

nonwhite_voters_average_above_rounded <- round(nonwhite_voters_average_above, 1)

})

```{r, echo=FALSE}

@@

#Here is the BELOW value

suppressMessages({nonwhite_voters_average_below <- inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_race_nonwhite_abovebelow_natl) %>% 
  summarise(average_below_nonwhite = mean(pct_r_2020)) %>% 
  filter(pct_race_nonwhite_abovebelow_natl=="BELOW") %>% 
  pull(average_below_nonwhite)

nonwhite_voters_average_below_rounded <- round(nonwhite_voters_average_below, 1)
})

@@

#Here is the ABOVE value

suppressMessages({average_pct_ed_college_all_above_natl <- inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise(average_above_college = mean(pct_r_2020)) %>% 
  filter (pct_ed_college_all_abovebelow_natl == "ABOVE") %>% 
  pull(average_above_college)

average_pct_ed_college_all_above_natl_rounded <-  round(average_pct_ed_college_all_above_natl, 1)

})

```{r, echo=FALSE}

#Here is the BELOW value

suppressMessages({average_pct_ed_college_all_below_natl <- inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise(average_below_college = mean(pct_r_2020)) %>% 
  filter (pct_ed_college_all_abovebelow_natl == "BELOW") %>% 
  pull(average_below_college)

average_pct_ed_college_all_below_natl_rounded <-  round(average_pct_ed_college_all_below_natl, 1)

})

# **`r nonwhite_voters_average_above_rounded`% of Trump's voters are part of the counties above national average for people who identify as nonwhite.**

In the aftermath of the 2020 Presidential campaign, both the results of the election and the demographics of the voter pool of each county were analyzed for both the Republican candidate Donald Trump and the Democratic candidate Joe Biden. Particularly, Trump's mean percentage of votes by U.S. counties was analyzed for both the number of nonwhite voters and the number of college graduates and if they were above or below the U.S. national average.

It comes as a surprise that more than `r nonwhite_voters_average_above_rounded` percent of Trump's voters are part of the counties above the national average for people who identify as nonwhite, while approximately `r nonwhite_voters_average_below_rounded` of votes are part of counties below the national average for non white-identified people. Furthermore, on average, `r average_pct_ed_college_all_above_natl_rounded` percent of Trump's votes come from counties with people above the national average for college graduates, versus `r average_pct_ed_college_all_below_natl_rounded` percent below the national average for college graduates.

These results come as a surprise since it is known that the Democratic party has the majority of college graduate and nonwhite votes. Although this is still the case, the Democratic party cannot solely rely on these metrics to ensure victory since more college graduates and nonwhite voters can be switching their minds and increasing their support for the Republican party.

This trend can be seen in states such as Texas. Some working-class Hispanics did not feel completely represented by the Democratic party in the last gubernatorial race in November, 2022. The Democratic candidate, Beto O'Rourke is a citizen of El Paso and has been called out by activists from the organization Sunrise El Paso for not supporting them in their fight against the privatization of their electric companies. The city of El Paso is not on the Texas electrical grid, allowing for the company J.P Morgan to privatize it and charge expensive bills. Representatives of Sunrise El Paso claim that O'Rourke is too much of a Central Democrat and does not want to compromise the funding of his campaign by selecting fights with big corporations. The midterm results showcased that O'Rourke lost to Republican candidate Greg Abbott, and that Abbott secured a good portion of votes from the Hispanic community.

### 2023-05-09 censusmap.qmd
another commit
final commit

```{r}
# Census API key
#Retrived from the API Link on their page 
census_api_key("75fa32dac9122cf2c43fb6e5d0ed1e7ead626185")

@@

# Join census data with spatial data
tx_map_data <- left_join(tx_counties, tx_census, by = c("NAME" = "NAME"))
```

>>>>

```{r}
# Census API key
#Retrived from the API Link on their page

census_api_key("75fa32dac9122cf2c43fb6e5d0ed1e7ead626185")

@@

### 2023-05-09 newsstory.qmd
final commit

# Voter Suppression in Texas

@@

In response to Abbott's actions, Trump wrote a letter claiming that Texans \"do not trust (Abbott's) election system, and they want (his) leadership on this issue, which is the number one thing they care about.\"

@@

\"Instead of working on issues that actually matter, like protecting school kids from Covid or fixing our failing electrical grid, Abbott is focused on rigging our elections and implementing extreme, right-wing policies,\" said O\'Rourke to The New York Times.

>>>>

# **Voter Suppression in Texas**

@@

In response to Abbott's actions, Trump wrote a letter claiming that Texans "do not trust (Abbott's) election system, and they want (his) leadership on this issue, which is the number one thing they care about."

@@

"Instead of working on issues that actually matter, like protecting school kids from Covid or fixing our failing electrical grid, Abbott is focused on rigging our elections and implementing extreme, right-wing policies," said O'Rourke to The New York Times.

### 2023-05-09 portfolio.qmd
final commit

## **Burberry: the brand\'s dilemma with sustainability**

@@

With the influx of these concerned consumers, brands must adjust their strategies and undertake the topic of sustainability to stay competitive. For example, in the luxury fashion industry, fur has always been an important medium. In recent years, there has been a decrease in demand for the luxurious animal by product because there is a heightened awareness of the detrimental effects of the fur industry. For that reason, some designers are beginning to swap fur for something faux. In my opinion, this shift is beneficial and will have a positive impact on a luxury brand\'s reputation.

One high end brand that is highly invested in sustainability is Burberry, the British fashion house created in 1865 by Thomas Burberry. The company proved to be unique in the high fashion market when Thomas created Gabardine, a \'breathable, weatherproof and hardwearing fabric revolutionising rainwear\'. Also adding to the originality of the brand is Thomas Burberry\'s creation of the iconic pattern-based scarves, purses, and trench coats. In 2017, when the company\'s annual report was released, environmental activists protested the brand because it had burned almost \$29 million dollars of unsold inventory in a year. With the major repercussion of the brand\'s lack of sustainability commitment, social media backlash, the movement #Burnberry arose to demand change. The protests proved to be effective since Burberry is now dedicated to improving sustainability inside the company.

As of now, the brand uses 58% of its energy and 68% of its electricity from clean energy sources and Burberry plans to stay committed to sustainability in the future. The brand has made several promises to become more sustainable for 2022. For instance, Burberry has committed to stopping the use of real fur and burning unwanted items, become carbon neutral by 2022, and operate entirely by renewable energy. In addition, Burberry has released a \'Responsibility Agenda\' to set goals for increasing sustainability within the company. The agenda focuses on three areas: product, company, and communities. Each area is centered on achieving progress by making the brand\'s product more recyclable, ridding the practice of burning unsaleable products, and impacting teachers and students by increasing mentoring programs in developing regions.

The \'Responsibility Agenda\' goals abides with the Paris Climate Agreement and the UN\'s 17 Sustainable Development Goals. Burberry committed to becoming a 100% renewable energy run company, and to create a Manufacturing Excellence Program focused on enhancing worker wellbeing in the company\'s supply chain. The brand has been recognized with two awards for its sustainable fashion approach: Leading luxury brand in the \'Textiles, Apparel & Luxury Goods\' sector by the DJSI 2018, and \'Bronze Class\' in RobecoSAM\'s 2018 Sustainability Yearbook.

@@

\
   Fashion is always trying to reinvent itself. During the pandemic, various fashion trends were revived. One in particular was psychedelic patterns. Inspired by the 70s, this type of print was first popularized by counterculture teenagers in the midst of social and political turmoil. The 'LSD fabric' made of colorful lines refers to a 'hallucinogenic\'' pattern that became the inspiration for other psychedelic shapes, and made it possible for the pattern to be produced quickly and appear as a 'groovy' style in movies. These patterns are known for their bold colors, sinuous lines and natural shapes.

@@

During the pandemic, people started to question the meaning of traditional fashion, and how can the industry be more inclusive. Quarantine gave people a lot of time to think about this aspect, giving more freedom to reinvent the industry and history of fashion shows. Rihanna was one of the designers who invested in this topic and brought to the public a fashion show that empowered people\'s sexuality and encouraged all to take ownership and pride of their bodies when using lingeries.

@@

Another concept incorporated in this show was a virtual theatrical experience. If you watched it, you know that the show was filled with dancers and performances. This unique model enhanced the powerfulness of the show when portraying diversity of genders, races, body heights, and weights. It was a very creative approach in comparison to regular fashion shows, especially when it is important to keep consumers\' attention to the show now that it has a virtual format. The performance is available on Amazon prime. It is shocking how a year ago this was never considered a possibility, and now, it's the new norm.

In a world where Victoria Secrets\' scandal showed that there is still low tolerance when it comes to diversifying lingerie models in the runways, Rihanna questions this status quo with the release of her 2020 Savage X Fenty collection, and does that by capturing its public with an incredible virtual experience!

>>>>

## **Burberry: the brand's dilemma with sustainability**

@@

With the influx of these concerned consumers, brands must adjust their strategies and undertake the topic of sustainability to stay competitive. For example, in the luxury fashion industry, fur has always been an important medium. In recent years, there has been a decrease in demand for the luxurious animal by product because there is a heightened awareness of the detrimental effects of the fur industry. For that reason, some designers are beginning to swap fur for something faux. In my opinion, this shift is beneficial and will have a positive impact on a luxury brand's reputation.

One high end brand that is highly invested in sustainability is Burberry, the British fashion house created in 1865 by Thomas Burberry. The company proved to be unique in the high fashion market when Thomas created Gabardine, a 'breathable, weatherproof and hardwearing fabric revolutionising rainwear'. Also adding to the originality of the brand is Thomas Burberry's creation of the iconic pattern-based scarves, purses, and trench coats. In 2017, when the company's annual report was released, environmental activists protested the brand because it had burned almost \$29 million dollars of unsold inventory in a year. With the major repercussion of the brand's lack of sustainability commitment, social media backlash, the movement #Burnberry arose to demand change. The protests proved to be effective since Burberry is now dedicated to improving sustainability inside the company.

As of now, the brand uses 58% of its energy and 68% of its electricity from clean energy sources and Burberry plans to stay committed to sustainability in the future. The brand has made several promises to become more sustainable for 2022. For instance, Burberry has committed to stopping the use of real fur and burning unwanted items, become carbon neutral by 2022, and operate entirely by renewable energy. In addition, Burberry has released a 'Responsibility Agenda' to set goals for increasing sustainability within the company. The agenda focuses on three areas: product, company, and communities. Each area is centered on achieving progress by making the brand's product more recyclable, ridding the practice of burning unsaleable products, and impacting teachers and students by increasing mentoring programs in developing regions.

The 'Responsibility Agenda' goals abides with the Paris Climate Agreement and the UN's 17 Sustainable Development Goals. Burberry committed to becoming a 100% renewable energy run company, and to create a Manufacturing Excellence Program focused on enhancing worker wellbeing in the company's supply chain. The brand has been recognized with two awards for its sustainable fashion approach: Leading luxury brand in the 'Textiles, Apparel & Luxury Goods' sector by the DJSI 2018, and 'Bronze Class' in RobecoSAM's 2018 Sustainability Yearbook.

@@

\
   Fashion is always trying to reinvent itself. During the pandemic, various fashion trends were revived. One in particular was psychedelic patterns. Inspired by the 70s, this type of print was first popularized by counterculture teenagers in the midst of social and political turmoil. The 'LSD fabric' made of colorful lines refers to a 'hallucinogenic'' pattern that became the inspiration for other psychedelic shapes, and made it possible for the pattern to be produced quickly and appear as a 'groovy' style in movies. These patterns are known for their bold colors, sinuous lines and natural shapes.

@@

During the pandemic, people started to question the meaning of traditional fashion, and how can the industry be more inclusive. Quarantine gave people a lot of time to think about this aspect, giving more freedom to reinvent the industry and history of fashion shows. Rihanna was one of the designers who invested in this topic and brought to the public a fashion show that empowered people's sexuality and encouraged all to take ownership and pride of their bodies when using lingeries.

@@

Another concept incorporated in this show was a virtual theatrical experience. If you watched it, you know that the show was filled with dancers and performances. This unique model enhanced the powerfulness of the show when portraying diversity of genders, races, body heights, and weights. It was a very creative approach in comparison to regular fashion shows, especially when it is important to keep consumers' attention to the show now that it has a virtual format. The performance is available on Amazon prime. It is shocking how a year ago this was never considered a possibility, and now, it's the new norm.

In a world where Victoria Secrets' scandal showed that there is still low tolerance when it comes to diversifying lingerie models in the runways, Rihanna questions this status quo with the release of her 2020 Savage X Fenty collection, and does that by capturing its public with an incredible virtual experience!

### 2023-05-09 walkthrough_R_concept.qmd
final commit

---
title: " Walkthrough"
format:
  html:
    self-contained: true
---

Here, I am going to walk you through the steps of loading your packages and dataframe, and using functions such as *filter, group_by, mutate and summarise*.

@@

```
Another good tool I can use is filtering out only the members who are in favor of the impeachment. Then, after filtering it, I can check how many won their 2018 election by less than 5 percentage points vs. more - which is found in the row `margin_flag_2018`.

@@

>>>>

---
title: " Walkthrough"
author: "Giovanna Lino"
format:
  html:
    self-contained: true
---

Here, I am going to walk you through the steps of loading your packages and dataframe, and using commands such as *filter, group_by, mutate and summarise*. Then, I will also walk you through the concept of *functions* and how to create them using certain data sets.

# **Filter, Group_by, Mutate & Summarise**

@@

Another good tool I can use is filtering out only the members who are in favor of the impeachment. Then, after filtering it, I can check how many won their 2018 election by less than 5 percentage points vs. more - which is found in the row `margin_flag_2018`.

@@

# **Functions**

Now, I am going to walk you through the concepts of functions and how to create them. First, I create a function that will return the sum of 2 numbers by assigning a name to my function. In this case, I assigned 'sum_two_numbers' to my function of number x and y. Then, I put 'return' to add both values. At last, I add the assign the name 'result' to my sum_two_numbers(5,7) and print the result.

sum_two_numbers <- function(x,y){

return(x + y)

}

result <- sum_two_numbers (5,7)
print(result)

Now, I can create a function that will return the mean of a list of numbers fed to it. I use the same step as above but instead of summing values, I am generating their mean. The 'list' portion refers to the numbers I assigned in the bottom of the code.

mean_of_numbers <- function(list){
  return(mean(list))
}

list <- c(1,4,6,8,10)
result <- mean_of_numbers(list)
print(result)

Now, I expand on the function I made above, and instead of just returning the mean number alone, I have it return a sentence. \*"The mean is \_\_\_"\*. (Where the blank is the mean.)

I followed the same steps as above by assigning a name to my function and including a list of numbers in the function. After, I just added the text I want generated by using the 'return' and the 'paste0' commands.

```{r}
mean_of_numbers_with_text <- function(list){
  return(paste0("The mean is ", mean_of_numbers(list)))
}

list <- c(1,4,6,8,10)
result <- mean_of_numbers(list)
print(result)

Another option I did was to create a code that showcases the mean with text. I added a list of numbers and generate its mean with the 'mean' command. With the 'paste0', I use the text I want printed and the mean I want from the respective list.

```{r}
numberlist <- c(12,14,15)

resultingmean <- mean(numberlist)

paste0("The mean is ", resultingmean)
```

Now, we'll use the `flights` dataset from the package `nycflights13`. It contains records of individual airline flights departing at New York's three metro airports, JFK, LaGuardia (LGA) and Newark (EWR). Run the following chunk to load the data.

```{r}
install.packages("nycflights13")
```

```{r}
#load the data to use for the rest of the assignment questions
flights <- nycflights13::flights

head(flights)
```

Using the `flights` dataset, I created a function that filters the data frame to only include records from a single originating airport (column name `origin`).

In other words, it should let you enter an origin airport and return all the records from just that airport. Then, I create the function I want by assigning a name to it and return the result. In this case I assigned the name 'filter_flights_by_origin'.

```{r}
#creating the function

airportname <- "LGA"

flights %>% 
  filter(origin == airportname)

filter_flights_by_origin <- function(airportname) {
  result <- flights %>% 
  filter(origin == airportname)

  return(result)
}
```

Now, I use the assigned name to only return the values from the origin I selected.

```{r}
#returning values

filter_flights_by_origin("LGA")

Now, using the `flights` data set, I created a function that:

First, filters the data frame to only include records from a single originating airport (column name `origin`)...

```{r}
airportname <- "LGA"

flights %>% 
  filter(origin == airportname) %>% 
  count(carrier)
```

... And then ,aggregates the results by airline (`carrier`) to show how many flights each airline has from that airport. Here, I used the filter, count, and function commands. I combined all of the steps above into one. A different command here is the 'head' command that shows the top value of the list since I included the #1 inside the brackets.

```{r}
count_byairline_for_one_airport <- function(airportname) {
  result <- flights %>% 
    filter(origin == airportname) %>% 
    count(carrier, sort = TRUE) %>% #sort = TRUE showcases from the highest value to the lowest 
    head(1) #for the top value of the list

  return(result)
}

count_byairline_for_one_airport("EWR")
